SIR Model With Testing Flows and Isolation

### 2020-11-16 SIR_manuscript.tex
Add files via upload
minor updates
clean up and organized
updated panel plot and manuscript

\documentclass{article}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsmath}
\bibliographystyle{apalike}

%%Packages and commands added by Fady
\usepackage{xcolor}
\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
\newcommand{\Rnum}{\mathcal{R}_0}
\usepackage{xspace}
\usepackage[nameinlink,capitalize]{cleveref}

@@

- Our work; we developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism. 
- The basic reproduction number was derived analytically as a threshold of the dynamics. 
- Our simulation shows ...

@@

% %%%%%%%
\section{Method}
\begin{itemize}
\item $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively. We assume that $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
\item $\omega$: the rate of onward flow from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.  It has units of 1/time.
\item $\gamma$: recovery rate (1/time).
\item $\rho$: per capita testing intensity across the whole population (1/time).
\item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
\item $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
\item $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
\end{itemize}

\includegraphics[width=\linewidth]{./pix/sir_comp.png}

The model is

\begin{align}
\label{model}
 d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
 d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n \\
 d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p \\
 d S_t/dt &= -\Lambda S_t + \omega S_p \\
 d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
 d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
 d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p \\
 d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t  \\
 d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
 d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
 d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
 d R_t/dt&= \gamma I_t + \omega R_p  \\
 dN/dt &= \omega (S_n + I_n + R_n)   \\
 dP/dt &= \omega(I_p + R_p).
\end{align}

@@

@@

@@

\subsection{Expensive vs. cheap tests}
\fady{needs to be written \dots}

@@

>>>>

\documentclass{article}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{xcolor}
\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
\newcommand{\Rnum}{\mathcal{R}_0}
\usepackage{xspace}
\usepackage[nameinlink,capitalize]{cleveref}

\bibliographystyle{apalike}

@@

- What did we do? Why novel? 
we developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism. 
- The basic reproduction number was derived analytically as a threshold of the dynamics. 
- Our simulation shows ...

@@

% %%%%%%%
\section{Method}

We used a SIR (Susceptible-Infectious-Recovered) deterministic model. In order to collect cumulative reported negative or positive tests, two 'integrator' or 'accumulator' compartments, $N$ and $P$, were incorporated. The flowchart is

\begin{figure}[!h] 
\begin{center} 
\includegraphics[scale=0.3]{./pix/sir_comp.png}
\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model. Say more here to make it stand on its own.
\label{fig:flowchart}} 
\end{center} 
\end{figure}

The model is
\begin{align}
\label{model}
 d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
 d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n \\
 d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p \\
 d S_t/dt &= -\Lambda S_t + \omega S_p \\
 d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
 d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
 d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p \\
 d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t  \\
 d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
 d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
 d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
 d R_t/dt&= \gamma I_t + \omega R_p  \\
 dN/dt &= \omega (S_n + I_n + R_n)   \\
 dP/dt &= \omega(I_p + R_p),
\end{align}

where $\beta$, (1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively, $\Lambda$ is the force of infection and defined as 
\begin{equation}
\label{Lambda}
\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},
\end{equation}
where $N_0$ is the total population size. Further, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population. Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$. The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing defined as \begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W}.
\end{equation}
Thus, the weighted testing rate defined as $F_Z=\sigma W_Z$, for $Z \in \{S,I,R\}$.

- In this model we assumed: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) for simplicity of the analysis, a perfectly specific test, $p_S=0$, was assumed, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals. 
-why not bith and death are included?

@@

The Disease-Free Equilibrium, DFE hearafter, for the above SIR model is given by the solution of the following system $S_u+S_n=N0$ and $F_S S_u-\omega S_n=0$. The DFE is

\begin{equation}
\label{dfe}
S_n^*= \frac{\rho}{\omega} N_0, S_u^*= N_0-S_n^*, \text{and} I_j=R_j=0 \ \forall j.
\end{equation}

The Basic reproduction number, $\Rnum$, was calculated by using the method of next generation matrix \cite{van2002reproduction}. See the Appendix for details.

\begin{equation}
\label{R0}
\Rnum= (A \times S_u^* + B \times S_n^*) \times C, 
\end{equation}
where
$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_random.png}
        \caption{}\label{fig:res1_a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
\caption{}\label{fig:res1_b}
\end{subfigure}
\end{figure}

@@

- "delay negative results" result

@@

\subsection{Expensive vs. cheap tests}
\fady{needs to be written \dots must discuss the fact that our analysis holds only when $\rho$ is relatively small, so if there is a test so cheap that $\rho$ does not stay small, our results do not apply.}

@@

% %%%%%%%
\section{supplementary material}

The next generation matrix, $G = F V^{-1}$, is in the lower triangular form with $F$ and $G$ as follows.

\begin{equation}\label{FV}
F= \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right],
 V=
 \left[ \begin {array}{cccc}  
F_I+\gamma&-\omega&0&0\\
-(1-p_I)F_I&\omega+\gamma&0&0\\
-p_I F_I&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right],
\end{equation}

\begin{equation}
G = \left[ \begin {array}{cc}
G_{11}&G_{12}\\
0&0
\end {array} \right], \text{ where } \\
G_{11} =C
\left[\begin {array}{cc}
A\,S_u & B\,S_u\\
A\,S_n & B\,S_n
\end {array}\right],
\end{equation}

The Eigenvalues of $G_{11}$, are 0 and $\Rnum$.

### 2020-11-16 codes/ali_test_sir.Rmd
updated panel plot and manuscript

```{r, out.width = "400px"}
knitr::include_graphics("../pix/sir_comp2.png")
```

@@

```{r packages, message=FALSE}
library(deSolve)
library(ggplot2)
library(tidyr)
library(McMasterPandemic)
library(directlabels)
unpack <- McMasterPandemic::unpack

@@

@@

```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
p1_temp <- (p1
    + geom_contour_filled(breaks=pretty(unique(df2$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
)
p1_temp
# filename = "R0contour_TTI.png"
# filename = "R0contour_random.png"
ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
```

@@

```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p11
    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(s~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
)
```

>>>>

```{r, out.width = "400px"}
knitr::include_graphics("../pix/sir_comp.png")
```

@@

```{r packages, message=FALSE}
library(deSolve)
library(ggplot2)
library(tidyr)
library(McMasterPandemic)
library(directlabels)
library(latex2exp)
unpack <- McMasterPandemic::unpack

@@

df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_t, NA, R0))

@@

```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
    theme(panel.spacing=grid::unit(0,"lines"))
p1_temp <- (p1
    + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
)
p1_temp
# ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")

@@

```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
p2_temp <- (p11
    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(s~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle("W[S]=1, W[I]=W[R]=1")
)
p2_temp
# ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
```
